Grove: A Convergent Collaborative Structure-Editor Calculus

### 2020-07-01 grove-paper.tex
Paper meta-data and some figures

%% Journal information
%% Supplied to authors by publisher for camera-ready submission;
%% use defaults for review submission.
\acmJournal{PACMPL}
\acmVolume{1}
\acmNumber{CONF} % CONF = POPL or ICFP or OOPSLA
\acmArticle{1}
\acmYear{2018}
\acmMonth{1}
\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}

@@

%% Bibliography style
\bibliographystyle{ACM-Reference-Format}
%% Citation style
%% Note: author/year citations are required for papers published as an
%% issue of PACMPL.
\citestyle{acmauthoryear}   %% For author/year citations

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Note: Authors migrating a paper from PACMPL format to traditional
%% SIGPLAN proceedings format must update the '\documentclass' and
%% topmatter commands above; see 'acmart-sigplanproc-template.tex'.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% Some recommended packages.
\usepackage{booktabs}   %% For formal tables:
                        %% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
                        %% http://ctan.org/pkg/subcaption

@@

%% Title information
\title[Short Title]{Grove: A Convergent Collaborative Structure-Editor Calculus}         %% [Short Title] is optional;
                                        %% when present, will be used in
                                        %% header instead of Full Title.
%\titlenote{with title note}             %% \titlenote is optional;
                                        %% can be repeated if necessary;
                                        %% contents suppressed with 'anonymous'
% \subtitle{Subtitle}                     %% \subtitle is optional
%\subtitlenote{with subtitle note}       %% \subtitlenote is optional;
                                        %% can be repeated if necessary;
                                        %% contents suppressed with 'anonymous'

%% Author information
%% Contents and number of authors suppressed with 'anonymous'.
%% Each author should be introduced by \author, followed by
%% \authornote (optional), \orcid (optional), \affiliation, and
%% \email.
%% An author may have multiple affiliations and/or emails; repeat the
%% appropriate command.
%% Many elements are not rendered, but should be provided for metadata
%% extraction tools.

%% Author with single affiliation.
\author{First1 Last1}
\authornote{with author1 note}          %% \authornote is optional;
                                        %% can be repeated if necessary
\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
\affiliation{
  \position{Position1}
  \department{Department1}              %% \department is recommended
  \institution{Institution1}            %% \institution is required
  \streetaddress{Street1 Address1}
  \city{City1}
  \state{State1}
  \postcode{Post-Code1}
  \country{Country1}                    %% \country is recommended
}
\email{first1.last1@inst1.edu}          %% \email is recommended

%% Author with two affiliations and emails.
\author{First2 Last2}
\authornote{with author2 note}          %% \authornote is optional;
                                        %% can be repeated if necessary
\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
\affiliation{
  \position{Position2a}
  \department{Department2a}             %% \department is recommended
  \institution{Institution2a}           %% \institution is required
  \streetaddress{Street2a Address2a}
  \city{City2a}
  \state{State2a}
  \postcode{Post-Code2a}
  \country{Country2a}                   %% \country is recommended
}
\email{first2.last2@inst2a.com}         %% \email is recommended
\affiliation{
  \position{Position2b}
  \department{Department2b}             %% \department is recommended
  \institution{Institution2b}           %% \institution is required
  \streetaddress{Street3b Address2b}
  \city{City2b}
  \state{State2b}
  \postcode{Post-Code2b}
  \country{Country2b}                   %% \country is recommended
}
\email{first2.last2@inst2b.org}         %% \email is recommended

@@

\subsection{Scenario A: Editing Different Parts of the Code}
A + B make edits and transmit to C, C is able to resolve without any assistance, 
order that the edits are received doesn't matter you get the same program state
at the end

\subsection{Scenario B: Edits to Nested Parts of the Code}
A edits part of the code

@@

@@

\subsection{Scenario D: Deletion of Code Being Edited}
B deletes code that A is editing (introduce notion of identity)

@@

\begin{theorem}[Commutativity (convergency?)]

@@

>>>>

%% Journal information
%% Supplied to authors by publisher for camera-ready submission;
%% use defaults for review submission.
\acmJournal{PACMPL}
\acmVolume{1}
\acmNumber{POPL} % CONF = POPL or ICFP or OOPSLA
\acmArticle{1}
\acmYear{2021}
\acmMonth{1}
\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}

@@

\bibliographystyle{ACM-Reference-Format}
\citestyle{acmauthoryear}

\renewcommand{\topfraction}{1} % Allow floats to take up the page
\renewcommand{\textfraction}{0}

%% Packages
\usepackage{booktabs}
\usepackage{subcaption}
\usepackage{tikz}

%%%%%%%%
% TikZ Stuff
%\usepackage{etex} % Fix "No room for new \dimen" error
\usepackage{shellesc} % Fix bug that breaks the tikz 'external' library
\usepackage{tikz}
\usetikzlibrary{babel} % Ensure compatibility the 'babel' package

\usetikzlibrary{external}
%\tikzexternalize % activate!
%\usepackage{lua-visual-debug}

\usetikzlibrary{arrows.meta} % Arrow Tips
\tikzset{>=Stealth}
%\tikzset{<=stealth}
%\tikzset{arrows={-Stealth[scale=50]}}
%\tikzset{edge from parent/.style={draw,->,line width=0.6pt}}
%\tikzset{wideline/.style={line width=0.7pt}}
%\tikzset{boldline/.style={color=black,line width=1.0pt}}

\usetikzlibrary{
  graphs,       % Graph *notation*
  graphdrawing, % Graph *layout*
}

\usegdlibrary{
  trees,
}

@@

%% Title information
\title[Grove]{Grove: A Convergent Collaborative Structure-Editor Calculus}
%\subtitle{Subtitle}

%% Author information
\author{Michael D. Adams}
\orcid{0000-0003-3160-6972}

\author{Eric Griffis}
\orcid{nnnn-nnnn-nnnn-nnnn}

\author{Cyrus Omar}
\orcid{0000-0003-4502-7971}
\affiliation{
  \position{Assistant Research Scientist}
  \department[0]{Computer Science and Engineering}
  \department[1]{Electrical Engineering and Computer Science}
  \department[2]{College of Engineering}
  \institution{University of Michigan}
  \streetaddress{Bob and Betty Beyster Building, 2260 Hayward Street}
  \city{Ann Arbor}
  \state{MI}
  \postcode{48109-2121}
  \country{USA}
}

@@

\subsection{User Interface}

TODO: include mocked-up screenshots

Say: "might not be a popup"

\subsection{Scenario A: Editing Different Parts of the Code}

\begin{figure}
\centering
TODO: commutativity diagram of screenshots with two cursors
\caption{TODO:Caption}
\label{fig:my_label}
\end{figure}

\begin{figure}
\centering
\begin{tikzpicture}
% TODO: texttt
% TODO: position the different graphs
% TODO: ballanced trees
% TODO: use tikz graph notation? tree notation?
\path (0cm,0cm) graph[tree layout] {
  + -> {
    3[as={\texttt{3}}], % TODO: put a cursor (green) here
    4 % TODO: put a cursor (red) here
  }
};
\path (-3cm,-2cm) graph[tree layout] {
  + -> {
    13, % TODO: put a cursor here
    4
  }
};
\path (3cm,-2cm) graph[tree layout] {
  + -> {
    3,
    14 % TODO: put a cursor here
  }
};
\path (0cm,-4cm) graph[tree layout] {
  + -> {
    13, % TODO: put a cursor here
    14 % TODO: put a cursor here
  }
};
\end{tikzpicture}
\caption{Scenario A: Editing Different Parts of the Code}
\label{fig:Scenario-A}
\end{figure}

A + B make edits and transmit to C, C is able to resolve without any assistance, 
order that the edits are received doesn't matter you get the same program state
at the end

\subsection{Scenario B: Edits to Nested Parts of the Code}

\begin{figure}
\centering
\begin{tikzpicture}
% TODO: texttt
% TODO: position the different graphs
% TODO: ballanced trees
% TODO: use tikz graph notation? tree notation?
\path (0cm,0cm) graph[tree layout] {
  lam[as={$\mathrm{\lambda}_{\texttt{x}}$}] -> {
    * -> {
      2,
      x
    }
  }
};
\path (-3cm,-2cm) graph[tree layout] {
  lam[as={$\mathrm{\lambda}_{\texttt{x}}$}] -> {
    + -> {
      1,
      * -> {
        2,
        x
      }
    }
  }
};
\path (3cm,-2cm) graph[tree layout] {
  lam[as={$\mathrm{\lambda}_{\texttt{x}}$}] -> {
    * -> {
      3,
      x
    }
  }
};
\path (0cm,-4cm) graph[tree layout] {
  lam[as={$\mathrm{\lambda}_{\texttt{x}}$}] -> {
    + -> {
      1,
      * -> {
        3,
        x
      }
    }
  }
};
\end{tikzpicture}
\caption{Scenario B: Edits to Nested Parts of the Code}
\label{fig:Schenario-B}
\end{figure}

A edits part of the code

@@

\begin{figure}
\centering
\begin{tikzpicture}
% TODO: texttt
% TODO: position the different graphs
% TODO: ballanced trees
% TODO: use tikz graph notation? tree notation?
\path (0cm,0cm) graph[tree layout] {
  lam[as={$\mathrm{\lambda}_{\texttt{x}}$}] -> {
    x
  }
};
\path (-3cm,-2cm) graph[tree layout] {
  lam[as={$\mathrm{\lambda}_{\texttt{x}}$}] -> {
    + -> {
      1,
      x
    }
  }
};
\path (3cm,-2cm) graph[tree layout] {
  lam[as={$\mathrm{\lambda}_{\texttt{x}}$}] -> {
    * -> {
      2,
      x
    }
  }
};
\path (2cm,-4cm) graph[tree layout] {
  lam[as={$\mathrm{\lambda}_{\texttt{x}}$}] -> {
    + -> {
      1,
      * -> {
        3,
        x
      }
    }
  }
};
\path (-2cm,-4cm) graph[tree layout] {
  lam[as={$\mathrm{\lambda}_{\texttt{x}}$}] -> {
    * -> {
      3,
      + -> {
        1,
        x
      }
    }
  }
};
\end{tikzpicture}
\caption{Scenario C: Direct Conflicts}
\label{fig:Scenario-C}
\end{figure}

DIAGRAM: lam x. x

DIAGRAM: lam x. 1 + x

DIAGRAM: lam x. 2 * x

DIAGRAM: Merge: diamond with "1 +" and "2 *"

DIAGRAM: Resolution \#1: lam x. 1 + 2 * x

DIAGRAM: Resolution \#2: lam x. 2 * (1 + x)

@@

\subsection{Scenario D: Deletion of Code Being Edited}

DIAGRAM: lam x. 2 + x

DIAGRAM: (lam x. HOLE) (2 + x)

DIAGRAM: lam x. 3 + x

DIAGRAM: Merge: (lam x. HOLE) (3 + x)

A deletes code that B is editing (introduce notion of identity)

@@

\begin{theorem}[Commutativity (convergence?)]

@@

We implemented GRV as a core OCaml library in approximately 1,500 lines of code, and a basic Web interface in approximately 600 lines of OCaml code and 250 lines of HTML+CSS+Javascript.

Compared to the core language, our implementation lacks support for lists.
